---
title: "Managing Felis catus impact on a protected island: Outcomes of an ongoing socio-ecological strategy in the Martín García Island Reserve (Buenos Aires, Argentina)"
author: "Ian Barbe, Lucía Inés Rodríguez-Planes, María del Rosario Jacoby, Andrea Szmelc, Gloria Domínguez, Nazareno Asín, María Eugenia Cueto, María Marcela Orozco"
email: iannbarbe@gmail.com
year: 2025
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r library}
library(tidyverse)
library(readxl)
library(lme4)
library(ggeffects)
library(performance)
library(emmeans) #comparaciones multiples e IC
library(writexl)
library(magrittr) #para pipe %>%
library(MuMIn) #comparacion de modelos con AIC
library(glmmTMB) #para modelado
library(MASS) #para binomial negativa en glm
library(pscl) #para correro Poisson con cero inflado
library(gridExtra) #fusion de graficos
library(cowplot) #fusion de graficos
```

```{r GLMM Felis catus}
## Cargado de tablas

setwd("E:/Análisis_R/IMG/Barbe_etal2025_MGI")

gatos <- read_excel("E:/Análisis_R/IMG/Barbe_etal2025_MGI\\data/gatos18-21.xlsx") %>%
  rename(abundancia = 1) %>%
  mutate(etapa = factor(etapa)) %>%
  mutate (tasa = abundancia/esfuerzo)%>%
  filter(zona == "Media")

#Para graficar lo que pasa en zona Baja ---
gatos_exp <- read_excel("E:/Análisis_R/IMG/Barbe_etal2025_MGI\\data/gatos18-21.xlsx") %>%
  rename(abundancia = 1) %>%
  mutate(etapa = factor(etapa)) %>%
  mutate (tasa = abundancia/esfuerzo)%>%
  filter(zona != "Alta") %>% 
  filter(zona == "Baja")

#Convertir etapa as character
gatos_exp$etapa <- as.character(gatos_exp$etapa)
class(gatos_exp$etapa)
gatos_exp$abundancia
```

## GLMM Felis catus plots


```{r GLMM Felis catus plots}

#### Grafico de gatos en BAJA----
gatos_baja_plot <- gatos_exp %>%
  ggplot(aes(x = etapa, 
             y = abundancia)) +
  geom_point(size=1.5) +
  geom_line(aes(group=abundancia))+
  labs(y = "Cat abundance", x = "Year", tag="",
       title="Low zone")+
  theme(      plot.tag.position = "topleft", 
              plot.title = element_text(size = rel(1.5),  face = "plain", hjust=0), 
              plot.title.position ="panel",
              axis.text=element_text(size=10,color="black"),
              axis.title=element_text(size=12,color="black"),
              panel.background=element_rect(color = "black"),
              strip.background=element_rect(color = "black"))+
         ylim(0,3)
gatos_baja_plot
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
