---
title: "Non-lethal management of Felis catus on a protected island: Outcomes of a successful ongoing socio-ecological strategy in Martín García Island Reserve, Argentina"
author: "Ian Barbe, Lucía Inés Rodríguez-Planes, María del Rosario Jacoby, Andrea Szmelc, Gloria Domínguez, Nazareno Asín, María Eugenia Cueto, María Marcela Orozco"
corresponding author email: marcelaorozco.vet@gmail.com
year: 2026
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

```{r library}
library(tidyverse)
library(readxl)
library(lme4)
library(ggeffects)
library(performance)
library(emmeans) 
library(writexl)
library(magrittr) 
library(MuMIn) 
library(glmmTMB) 
library(MASS) 
library(pscl)
library(gridExtra)
```

```{r GLMM Sighted cats}

#Medium zone
cats <- read_excel("R data/Fig3a,b.SightedCats18-21_GLMM.xlsx") %>%
  rename(abundance = 1) %>%
  mutate(year = factor(year)) %>%
  mutate (index = abundance/effort)%>%
  filter(zone == "Medium")

#Low zone ---
cats_exp <- read_excel("R data/Fig3a,b.SightedCats18-21_GLMM.xlsx") %>%
  rename(abundance = 1) %>%
  mutate(year = factor(year)) %>%
  mutate (index = abundance/effort)%>%
  filter(zone != "High") %>% 
  filter(zone == "Low")

#year as character
cats_exp$year <- as.character(cats_exp$year)
class(cats_exp$year)
cats_exp$abundance

<<<<<<< HEAD
=======
#### GLMM medium zone model for sigthed cats----

mod_cats <- glmer(abundance ~ year + (1|transect) + offset(log(effort)),
                   family = "poisson", data = cats)

check_overdispersion(mod_cats)
check_model(mod_cats)
summary(mod_cats)

## Likelihood Ratio Test
mod_cats %>% 
  update(.~.-year) %>% 
  anova(mod_cats, test = "Chisq") #AIC

#predicted values
cats$predicted <- predict(mod_cats, type= "response") 
cats$pred_index <- cats$predicted/cats$effort

#predicted per year
year_predicted <- predict(mod_cats, type="response")
year_predicted

#random predicted values with CI 
cats_predict_random <- ggpredict(mod_cats, terms = "year", type = "random")
cats_predict_random
```
```{r Tukey}
mod_cats %>% 
  emmeans(list(pairwise~year), type="response"   )
  
# CI
  emms_cats <- emmeans(mod_cats, "year", type="response")  
  summary(emms_cats)
  confint(pairs(emms_cats))
>>>>>>> 618f40bf3e2ed08b1e34710d8dd0d1286fb9ad07
```

```{r GLMM LIBRERIA}
library(lme4)
library(AICcmodavg)
library(ggeffects)
library(emmeans)
library(performance)
```
```{r AIC rankings}
#### GLMM medium zone model for sigthed cats----

# Null model
m0 <- glmer(
  abundance ~ offset(log(effort)) + (1 | transect),
  family = poisson,
  data = cats
)

# Year as fix factor
m1 <- glmer(
  abundance ~ year + offset(log(effort)) + (1 | transect),
  family = poisson,
  data = cats
)
```

```{r model selection}
cand_models <- list(m0, m1)
mod_names <- c("Null", "Year")

aictab(
  cand.set = cand_models,
  modnames = mod_names,
  second.ord = TRUE
)

aictab

```
```{r best model and predictions}
best_cat_model <- m1
summary(best_cat_model)

cats_pred <- ggpredict(
  best_cat_model,
  terms = "year",
  type = "random"
)

cats_pred

```
```{r emmeans}

emms_cats <- emmeans(
  best_cat_model,
  ~ year,
  type = "response"
)

# Predicted means + 95% CI
summary(emms_cats)

# Pairwise comparisons between years (Tukey)
pairs(emms_cats, adjust = "tukey")
confint(pairs(emms_cats, adjust = "tukey"))

```
```{r best model}
#Chequeos
check_overdispersion(best_cat_model)
check_model(best_cat_model)
```

```
### L. wiegmannii lizard -----
<<<<<<< HEAD
```{lizard GLM}
#model---
l.w_lizard <- read_excel("Fig4.Lw_lizard18-21_GLM.xlsx") %>%
=======
```{r lizard GLM}
l.w_lizard <- read_excel("R data/Fig4a.Lw_lizard18-21_GLM.xlsx") %>%
>>>>>>> 618f40bf3e2ed08b1e34710d8dd0d1286fb9ad07
  rename(abundance = 1) %>%
  mutate(year = factor(year)) %>%
  mutate(index = abundance/effort)

#model
mod_lizard1 <- glm(abundance ~ year +  offset(log(effort)),
               family = "poisson", data = l.w_lizard)


check_overdispersion(mod_lizard1)
check_model(mod_lizard1)

#'[Binomial negativa'] 
mod_lizard2 <- glmmTMB(abundance ~ year,data=l.w_lizard,
                          family="nbinom2") 

check_overdispersion(mod_lizard2)
check_model(mod_lizard2)

#AIC
AIC(mod_lizard1,mod_lizard2)

summary(mod_lizard2)
l.w_lizard$predicted <- predict(mod_lizard2, type= "response") 
lizard_predindex <- l.w_lizard$predicted/l.w_lizard$effort

lizard_predict <- ggpredict(mod_lizard2, terms = "year", type = "fixed")
lizard_predict

```
```{r Tukey}
## emmeans
mod_lizard2 %>% 
  emmeans(list(pairwise~year), type = "response")

#CI
emms_lizard <- emmeans(mod_lizard2, "year", type="response")
summary(emms_lizard)
confint(pairs(emms_lizard))
```

### Caprimulgus spp.-----
```{r nightjars GLM}
<<<<<<< HEAD
### Caprimulgus spp.-----
nightjars <- read_excel("Fig4b.Nightjars18-21_GLM.xlsx") %>% 
=======
nightjars <- read_excel("R data/Fig4b.Nightjars18-21_GLM.xlsx") %>% 
>>>>>>> 618f40bf3e2ed08b1e34710d8dd0d1286fb9ad07
  rename(abundance = 1) %>%
  mutate(year = factor(year))

#model
mod_ng <- glm(abundance ~ year +  offset(log(effort)),
                      family = "poisson", data = nightjars)
check_overdispersion(mod_ng)
check_model(mod_ng)

summary(mod_ng)

nightjars$predicted <- predict(mod_ng, type= "response") 
nightjars_predindex <- nightjars$predicted/nightjars$effort

# predicted values
nightjars_predict <- ggpredict(mod_ng, terms = "year", type = "fixed")
nightjars_predict

```
```{r Tukey}
## emmeans
mod_ng %>% 
  emmeans(list(pairwise~year), type = "response")
#CI
emms_ng <- emmeans(mod_ng, "year", type="response")
summary(emms_ng) 
confint(pairs(emms_ng))
```

<<<<<<< HEAD

####Tegu -----------
```{r }

# Tegu lizard GLMM---
library(lme4)
library(AICcmodavg)
library(emmeans)
library(performance)
library(dplyr)
library(readxl)
```


```{r modelos GLMM Tegu lizard}
=======
```{r GLMM Tegu lizard}
>>>>>>> 618f40bf3e2ed08b1e34710d8dd0d1286fb9ad07
tegu <- read_excel("R data/Fig5.Tegu18-21_GLMM.xlsx")%>%
  rename(abundance = 1) %>%
  mutate(year = factor(year))

##GLMM model
m_full <- glmer(
  abundance ~ year * zone + (1 | transect) + offset(log(effort)),
  family = poisson,
  data = tegu
)

m_add <- glmer(
  abundance ~ year + zone + (1 | transect) + offset(log(effort)),
  family = poisson,
  data = tegu
)

m_year <- glmer(
  abundance ~ year + (1 | transect) + offset(log(effort)),
  family = poisson,
  data = tegu
)

m_zone <- glmer(
  abundance ~ zone + (1 | transect) + offset(log(effort)),
  family = poisson,
  data = tegu
)

m_null <- glmer(
  abundance ~ 1 + (1 | transect) + offset(log(effort)),
  family = poisson,
  data = tegu
)

```

```{r comparo modelos}
models <- list(
  full = m_full,
  additive = m_add,
  year = m_year,
  zone = m_zone,
  null = m_null
)

aictab(
  cand.set = models,
  modnames = names(models)
)

```
```{r selecciono mejor modelo}

best_tegu <- m_add   # suponiendo que este fue el mejor

check_overdispersion(best_tegu)
summary(best_tegu)

```

```{r AÑO estimacion de efectos RR}
emms_tegu_year <- emmeans(best_tegu, ~ year, type = "response")
summary(emms_tegu_year)

pairs(emms_tegu_year, type = "response")
confint(pairs(emms_tegu_year))
```

```{r ZONA estimacion de efectos RR}
emms_tegu_zone <- emmeans(best_tegu, ~ zone, type = "response")
summary(emms_tegu_zone)

pairs(emms_tegu_zone, type = "response")
confint(pairs(emms_tegu_zone))

```
############### Interviews --------------------------
```{r Table5 interviews project need to continue}
#Table 5. Need to continue the project data ----------------

#Frequency of affirmative responses related to project continuation on MGI
#Table5 matrix (data from Table5_data.xlsx)
Table5 <- matrix(c(24, 2, 50, 1), nrow = 2, byrow = TRUE)
#chi-square test
projectchisq <- prop.test(Table5)
print(projectchisq)
```
```{r Fig6. cat abundance perceptons interviews}
#Fig.6. Community perceptions of cat abundance at the beginning of the project, by category (Surrounding houses and Island) and year (2019-2021)-------------

#island (data from Fig6. Island data.xlsx) -------------------------------
#2019 
island_2019 <- matrix(c(11, 3, 14), nrow = 1, byrow = TRUE)
colnames(island_2019) <- c("lower", "higher", "total")
# lower vs. higher
result_chi_cuad_island_2019 <- chisq.test(island_2019)
result_chi_cuad_island_2019$expected
print(result_chi_cuad_island_2019)

#island 2021
island_2021 <- matrix(c(20, 7, 27), nrow = 1, byrow = TRUE)
colnames(island_2021) <- c("lower", "higher", "total")
# lower vs. higher
result_chi_cuad_island_2021 <- chisq.test(island_2021)
result_chi_cuad_island_2021$expected
print(result_chi_cuad_island_2021)

#Surrounding houses (data from Fig6. Surrounding houses data.xlsx) ------------------
#2019
surroundings_2019 <- matrix(c(18, 2, 20), nrow = 1, byrow = TRUE)
colnames(surroundings_2019) <- c("lower", "higher", "total")
# lower vs. higher
result_chi_cuad_surroundings_2019 <- chisq.test(surroundings_2019)
print(result_chi_cuad_surroundings_2019)

#2021
surroundings_2021 <- matrix(c(26, 5, 31), nrow = 1, byrow = TRUE)
colnames(surroundings_2021) <- c("lower", "higher", "total")
# lower vs. higher
result_chi_cuad_surroundings_2021 <- chisq.test(surroundings_2021)
print(result_chi_cuad_surroundings_2021)
```